Wildcard: Spreadsheet-Driven Customization of Web Applications

### 2020-01-10 paper.tex
finished pdf version of demos

To get a sense of how it might feel to use Wildcard, let's see an
example of someone using it to help with booking a trip on the travel
sites Airbnb and Expedia. These demos are best viewed in the online
version of this paper, which includes videos:
\url{https://geoffreylitt.com/wildcard}

\begin{figure}
\hypertarget{fig:airbnb-demo}{%
\centering
\includegraphics{media/airbnb-demo-300dpi.png}
\caption{Using Wildcard to augment the Airbnb page for searching travel
accommodations}\label{fig:airbnb-demo}
}
\end{figure}

The user starts by opening up the Airbnb search listings page to look
for a place to stay. The page looks nice and mostly works well, but is
missing some key features. As mentioned before, this page doesn't allow
the user to sort by price. It also doesn't let them filter by user
rating. Using Wildcard, the user can add these small features, while
leaving the page's design and the rest of its functionality unchanged
(shown in fig.~\ref{fig:sorting}).

After manipulating the data, the user can close the table view and
continue using the website with its original visual design. The table
view offers a way to change the data backing a page, but does not need
to replace the original interface entirely.

@@

For example, it's tedious on Airbnb to click on listings one by one to
add them to a list of favorites. Using Wildcard, the user can just
select multiple rows and favorite all of them with one click. Similarly,
we can also open the detailed pages for many listings in new tabs.
(Shown in fig.~\ref{fig:actions})

@@

@@

\begin{Shaded}
\begin{Highlighting}[]
  \OperatorTok{\{}
\NormalTok{    fieldName}\OperatorTok{:} \StringTok{"name"}\OperatorTok{,} \CommentTok{// The name of the data field}
\NormalTok{    readOnly}\OperatorTok{:} \KeywordTok{true}\OperatorTok{,}    \CommentTok{// Whether the user can edit the field}
\NormalTok{    type}\OperatorTok{:} \StringTok{"text"}\OperatorTok{,}      \CommentTok{// The type of the field}
    \CommentTok{// Given the DOM element for the entire listing,}
    \CommentTok{// return the DOM element representing this field}
\NormalTok{    el}\OperatorTok{:}\NormalTok{ (row) }\KeywordTok{=>} \VariableTok{row}\OperatorTok{.}\FunctionTok{querySelector}\NormalTok{(}\VerbatimStringTok{\textasciigrave{}.}\SpecialCharTok{$\{}\NormalTok{titleClass}\SpecialCharTok{\}}\VerbatimStringTok{\textasciigrave{}}\NormalTok{)}\OperatorTok{,}
  \OperatorTok{\}}
\end{Highlighting}
\end{Shaded}

@@

The site adapter also needs to support the reverse direction: sending
updates from the table to the original page. Most DOM manipulation is
not performed directly by the site adapter; instead, the adapter
specifies how to find the divs representing data rows, and the core
platform mutates the DOM to reflect the table state. The only exception
is row actions (like favoriting an Airbnb listing), which are
implemented as imperative Javascript functions that can can mutate the
DOM, simulate clicks on buttons, etc.

>>>>

To get a sense of how it might feel to use Wildcard, let's see an
example of someone using it to help with booking a trip on the travel
sites Airbnb and Expedia.{ These demos are best viewed as videos in the
online version of this paper
(\url{https://www.geoffreylitt.com/wildcard}).}

\begin{figure}
\hypertarget{fig:airbnb-demo}{%
\centering
\includegraphics{media/airbnb-demo-300dpi.png}
\caption{Using Wildcard to augment the Airbnb search page for booking
accommodations}\label{fig:airbnb-demo}
}
\end{figure}

The user starts by opening up the Airbnb search listings page to look
for a place to stay. The page looks nice and mostly works well, but is
missing some key features. As mentioned before, this page doesn't allow
the user to sort by price. It also doesn't let them filter by user
rating. Using Wildcard, the user can add these small features, while
leaving the page's design and the rest of its functionality unchanged.{
fig.~\ref{fig:airbnb-demo} shows an overview of augmenting the Airbnb
site.}

First, the user opens up the Wildcard panel, which shows a table
corresponding to the search results in the page. As they click around in
the table, the corresponding row in the page is highlighted so they can
see the connection between the views.

Then, the user can use standard spreadsheet column header features to
sort the page by price and filter by rating{
(fig.~\ref{fig:airbnb-demo}B)}.

After manipulating the data, the user was able to close the table view
and continue using the website with its original visual design. The
table view offers a way to change the data backing a page, but does not
need to replace the original interface entirely.

@@

For example, it's tedious on Airbnb to click on listings one by one to
add them to a list of favorites. Using Wildcard, the user can just
select multiple rows and favorite all of them with one click{
(fig.~\ref{fig:airbnb-demo}D)}. Similarly, we can also open the detailed
pages for many listings in new tabs.

Now the user would like to jot down some notes on the pros and cons of
each listing. To do this, they can simply type notes into an additional
column next to each listing, and the notes appear inside the listings in
the original UI{ (fig.~\ref{fig:airbnb-demo}A)}. These annotations are
also persisted in the browser for future retrieval.

Wildcard also includes a formula language which enables more
sophisticated tweaks that fetch external data and perform computations.

When traveling without a car, it's nice to evaluate potential places to
stay based on how walkable the surroundings. Using Wildcard formulas, we
can integrate Airbnb with Walkscore, an API that rates the walkability
of any location on a 1-100 scale. When we call the \texttt{WALKSCORE}
formula with the latitude and longitude of the listing, it returns the
walk score as the cell value. Because the cell's contents are injected
into the page, the score also shows up in the page body{
(fig.~\ref{fig:airbnb-demo}C)}.

@@

It might seem that Wildcard is only useful on websites that display
lists of tabular data like search results. But in fact, the table
metaphor is flexible enough to represent many types of data. For
example, a flight search form on Expedia can be represented as a single
row, with a column corresponding to each input. {
fig.~\ref{fig:expedia-demo} shows an overview of augmenting the Expedia
site.}

In previous examples the data extracted from the site was marked as
read-only; users cannot change the name or price of an Airbnb listing.
In this next case, the cells are marked as writable, so that changes in
the table are reflected in the form inputs. This becomes useful when
combined with GUI widgets for editing the value of a table cell.

Filling in dates for a flight search typically requires opening up a
separate calendar app to find the right dates, and then manually copying
them into the form. In Wildcard, we can make this easier by providing a
datepicker widget that has privileged access to the user's calendar
information{ (fig.~\ref{fig:expedia-demo}B)}.

@@

\begin{Shaded}
\begin{Highlighting}[]
\OperatorTok{\{}
\NormalTok{  fieldName}\OperatorTok{:} \StringTok{"name"}\OperatorTok{,} \CommentTok{// The name of the field}
\NormalTok{  type}\OperatorTok{:} \StringTok{"text"}\OperatorTok{,}      \CommentTok{// The type of the field}
\NormalTok{  readOnly}\OperatorTok{:} \KeywordTok{true}\OperatorTok{,}    \CommentTok{// Whether the user can edit the field}
\NormalTok{  el}\OperatorTok{:}\NormalTok{ (row) }\KeywordTok{=>} \VariableTok{row}\OperatorTok{.}\FunctionTok{querySelector}\NormalTok{(}\VerbatimStringTok{\textasciigrave{}.}\SpecialCharTok{$\{}\NormalTok{titleClass}\SpecialCharTok{\}}\VerbatimStringTok{\textasciigrave{}}\NormalTok{)}\OperatorTok{,} \CommentTok{// Extract DOM element}
\OperatorTok{\}}
\end{Highlighting}
\end{Shaded}

@@

The site adapter also needs to support the reverse direction: sending
updates from the table to the original page. Most DOM manipulation is
not performed directly by the site adapter; instead, the adapter
specifies how to find the spans representing data rows, and the core
platform mutates the DOM to reflect the table state. The only exception
is row actions (like favoriting an Airbnb listing), which are
implemented as imperative Javascript functions that can can mutate the
DOM, simulate clicks on buttons, etc.
